A Triple Origin for Twin Blue Stragglers in Close Binaries

### 2019-02-24 BSS7782.tex
Add files via upload

We propose a formation mechanism for twin blue stragglers (BSs) in
compact binaries that involves mass transfer from an evolved outer
tertiary companion on to the inner binary via a circumbinary disk.  We
apply this scenario to the observed double BS binary WOCS 7782 in
the old open cluster NGC 188, and show that its observed properties
are naturally reproduced within the context of the proposed model.
Based on this model, we predict the following properties for twin BSs:
(1) For the outer tertiary orbit, the initial orbital period should
lie between 220 days $\lesssim$ P$_{\rm out}$ $\lesssim$ 1100 days,
assuming initial masses for the inner binary components of $m_{\rm 1}
= 1.1$ M$_{\odot}$ and $m_{\rm 2} =$ 0.9 M$_{\odot}$ and an outer
tertiary mass of $m_{\rm 3} = 1.4$ M$_{\odot}$.  \textbf{(2)} After Roche-lobe
overflow, the outer star turns into a white dwarf (WD) of mass 0.43 to
0.54\,\MSun. There is a correlation between the mass of this WD and
the outer orbital period: more massive WDs will be on wider orbits.
(3) The rotational axes of both BSs will be aligned with each other
and the orbital plane of the outer tertiary WD. (4) The BSs will have
roughly equal masses, independent of their initial masses (since the
lower mass star accretes the most).  The dominant accretor should,
therefore, be enriched more effectively by the accreted material.  As
a result, one of the BSs will appear to be more enriched by either He,
C and O or by s-process elements, depending on if the donor started to
overflow its Roche lobe on, respectively, the red giant or asymptotic
giant branch.  (5) Relative to old dense clusters with high-velocity
dispersions, twin BSs in close binaries formed from the proposed
mechanism should be more frequent in the Galactic field and younger
open clusters with ages $\lesssim$ 4-6 Gyr, since then the donor will
have a radiative envelope. (6) the orbit of the binary BS will have a
small semi-major axis (typically $\aplt 0.3$\,au) and be close to
circular ($e \aplt 0.2$).

@@

Even more curious, there exists in the old OC NGC 188 a
double BS binary, called WOCS 7782 \citep{geller08}.  The BS
population in NGC 188 has a bi-modal period-eccentricity distribution.
As discussed in \citet{2011MNRAS.410.2370L}, this could be hinting at
a triple origin for at least some subset of the total BS population.
As for WOCS 7782, \citet{2009Natur.462.1032M} observed a compact and
mildly eccentric (i.e., $e \sim 0.1$) binary star system with an
orbital period of $\sim$ 10 days hosting two roughly equal-mass BSs.  During a given binary-binary interaction, the probability
that not one but two direct (MS-MS) collisions will occur is less than
$10^{-2}$
\citep{1989AJ.....98..217L,2011MNRAS.410.2370L,2012MNRAS.425.2369L}.
Plus, binaries with collision products typically have relatively long
orbital periods \cite{2011Sci...334.1380F}. Dynamically, it is
difficult to form a short-period binary composed of two collision
products during a collisional interaction in a star cluster
\citep{2011MNRAS.410.2370L,2011Sci...334.1380F}.  So, how did WOCS 7782 form?

@@

In our scenario, we start with a binary star with component masses
$m_{\rm 1}$ and $m_{\rm 2}$ that is orbited by a tertiary of mass
$m_3$. The inner and outer binary orbital semi-major axes are denoted
a$_{\rm in}$ and a$_{\rm out}$, respectively.  For clarity we 
assume both orbits, the inner as well as the outer, to have negligible
eccentricity and low inclination.  These assumptions are also
supported by the population of observed low-mass triples
\citep{2010yCat..73890925T,2018ApJ...854...44M}.  This initial configuration for our
assumed formation scenario for WOCS 7782, described below, is
depicted in figure~\ref{fig:fig1}.

@@

Figure~\ref{fig:fig2} shows the parameter space in the P$_{\rm
  out}$-P$_{\rm in}$-plane for WOCS 7782.  We assume initial
component masses of $m_{\rm 1} = 1.1$ M$_{\rm \odot}$ and $m_{\rm 2} =
0.9$ M$_{\rm \odot}$ for the inner binary components, and $m_{\rm 3} =
1.4 $M$_{\rm \odot}$ for the outer tertiary.  We compare the
circularization radius to the semi-major axis of the inner binary, for
which we require $a_{\rm c} > a_{\rm in}$, after folding in all
constraints from the requirements for dynamical stability (listed in
the caption of figure~\ref{fig:fig2}), and the assumption of an outer tertiary that is Roche
lobe-filling.  
%\simon{Why do you consider it important to mention that the orbit is hard, and for what environment did you consider this? - NL: Deleted.}  
Note that the range of
plotted orbital periods P$_{\rm in}$ corresponding to a contact state
for the inner binary lies outside the range of plotted values for
$P_{\rm in}$ (for components with radii of 1 R$_{\odot}$),
since it does not contribute to constraining the outer orbital
properties.  The thick horizontal solid red line shows the allowed
range of outer semi-major axes, after folding in all of 
the aforementioned criteria.  These constraints result in a rather narrow
range of initial conditions for the outer orbit, namely 2.2 $\times$
10$^{2}$ days $\le$ P$_{\rm out}$ $\le$ 1.1 $\times$ 10$^3$ days, \textit{which also directly translate into constraints on the final outer tertiary orbit.}
%, for
%which the tertiary star overfills its Roche lobe, the triple is (at
%least initially) dynamically stable and the accretion stream forms a
%gaseous ring around the inner binary.

\textbf{Finally, we compute the timescales for our hypothesized triple to undergo a direct interaction with another single or binary star.  Using the same assumptions for the host cluster properties of NGC 188 outlined in Section 3.2 of \citet{leigh11} (right-hand column), we find upon setting the single-triple (1+3) and binary-triple (2+3) timescales equal to the expected duration of the mass transfer phase (i.e., $\sim$ 1 Myr) critical outer orbital periods for triples of, respectively, 8.1 $\times$ 10$^8$ and 2.4 $\times$ 10$^9$ days.  These critical outer tertiary orbital periods, which correspond to the times for a \textit{specific} triple to undergo an interaction, correspond to 1+3 and 2+3 interaction times that are much longer than the maximum predicted outer orbital period of the hypothesized white dwarf tertiary in our scenario.  Hence, we do not expect the mass transfer process to be interrupted by either 1+3 and 2+3 interactions.}

@@

With the adopted parameters, we can also estimate the final orbital
period of the left-over core from the tertiary star after 
mass transfer.  The change in orbital separation due to
non-conservative mass transfer can be expressed in terms of the mass
of the outer star before and after mass transfer, i.e. $m_3$ and $m'_3$
respectively, the total mass in the inner binary before ($m_{\rm in}$)
and after accretion ($m'_{\rm in}$) and the amount of angular momentum
lost per unit mass $\eta \simeq 3$. Adopting the relation between the
orbital separation before mass transfer ($a$) and after mass transfer
($a'$) from \cite{1995A&A...296..691P}
\begin{equation}
  {a' \over a} = \left( {m_3 m_{\rm in} \over m'_3 m'_{\rm in}} \right)^{-2}
  \left( {m_3 + m_{\rm in} \over m'_3 + m'_{\rm in}} \right)^{2\eta + 1},
\end{equation}
we arrive at the top brown curve in
figure\,\ref{fig:tertiarymass_vs_size}. This curve provides a prediction for 
the current orbital separation of the WD around the twin
BS 7782. Tidal effects during mass tranfer
have probably circularized the orbit, although some slight eccentricity
due to turbulent motion in the outer layers of the donor star may have
induced a small $e \ll 0.1$ eccentricity.

@@

We adopt initial masses of $m_{\rm 1} = 1.1$ M$_{\rm \odot}$ and
$m_{\rm 2} = 0.9$ M$_{\rm \odot}$ for the inner binary components, and
between $m_{\rm 3} = 1.2$ and $m_{\rm 3} = 1.4$ M$_{\rm \odot}$ for
the tertiary star.  We evolve the tertiary star using the MESA
stellar-evolution code \cite{2011ApJS..192....3P} to a radius of about
100\,\RSun\, and 150\,\RSun, at which point we assume it to overfill
it's Roche lobe (see red square in
figure~\ref{fig:tertiarymass_vs_size}).  We perform calculations for an
inner orbital separation of $a_{\rm in} = 0.10$\,au, $a_{\rm in} =
0.15$\,au and $a_{\rm in} = 0.20$\,au.  In total we performed 12
calculations at a resolution of 40k SPH particles and 12 at 80k.

@@

With these parameters the orbital separation of the outer binary
becomes $\sim 250$\,\RSun for the 100\,\RSun donor star and about
430\,\RSun\, for the more evolved donor star.  We adopt the outer
orbit to be circular and in the plane of the inner binary.
%% We
%% performed several additional calculations for inclined orbits (choosing
%% a relative initial inclination $i = 9^\circ$, $i = 45^\circ$, $i =
%% 90^\circ$ and $i = 180^\circ$, but we wi
%% The low inclination is motivated by observed stellar triples composed
%% of low-mass stars to favour co-planar configurations
%% \citep{2018ApJ...854...44M}.
%% To test the effect of highly inclined
%% and retrogade orbits we include the $i = 90^\circ$ and $i = 180^\circ$
%% cases.
%%SPZ: I removed the notion of low-inclination, because we already said as such in the intro.
In figure~\ref{fig:topview_at_t0} we present a top view of the initial
conditions for one of these calculations.

@@

To test the hypotheses that (1) the secondary in the inner binary
accretes more effectively than the primary star and to measure the
change to the inner orbit due to the Roche-lobe overflow of the outer
star, we perform a series of calculations in which we take the self
gravity and the hydrodynamical effects of the triple into account.
The results of these simulations are presented in
figure~\ref{fig:topview_at_t1000day} and
figure~\ref{fig:mass_vs_semimajor_axis}.  The first figure
(figure~\ref{fig:topview_at_t1000day}) shows the top view of the same
initial realization for which we presented the initial conditions in
figure~\ref{fig:topview_at_t0} but now at an age of 1091 days after
the onset of mass transfer. We add, to the left panel,
the equipotential surfaces in the orbital plane.

It is apparent that the mass transfer in the adopted triples leads to
a rather untidy evolution, since much of the donor mass is lost
through the second Lagrangian point to the right side of the donor
star in figure~\ref{fig:topview_at_t1000day}. A considerable amount of
mass is also lost through the third Lagrangian point (to the left of
the inner binary), although it is hard to actually quantize the amount
of material los, becuase an appreciable fraction is expected to rain
back onto the triple system.  One remaining question is how much mass
is eventually ejected altogether from the triple system and is
therefore not accreted to any of the two inner stars. This value is
hard to estimate from the simulations, but an accretion efficiency of
$\apgt 0.6$ is necessary to make the scenario feasible. Over the time
scale for which we performed the calculations, this efficiency is
reached, but it is not clear how the system responds at later stages.

@@

In the right-hand panel in figure~\ref{fig:mass_vs_semimajor_axis} we
present the evolution of the orbit for the 1.4\,\MSun\, donor for
several initial orbits of the inner binary. A more massive donor
appears to be more effective in producing a twin BS with parameters
consistent with the observed system 7782. There is more mass available
in the envelope of the donor star, and the orbital evolution of the
inner binary matches better with the anticipated evolution.  A more
massive donor may therefore have a lower accretion efficiency while
still accomodating the observed constraints.  The longer thermal time
scale of the stellar envelope of the higher-mass donor at the same
stellar radius eventually leads to a higher mass-transfer rate, and
therefore to a lower accretion efficiency. However, the larger mass
budget in the envelope appears to compensate.

The orbit of the inner binary expands in these cases as a result of
accretion onto the inner two stars. All three cases for the
1.4\,\MSun\, donor presented in
figure~\ref{fig:mass_vs_semimajor_axis} the inner orbit expands at
about the same rate. Consequently, the inner binaries that start with
$a = 0.15$\,au and $a=0.20$\,au eventually become dynamically
unstable.  The binary with an initial separation of $0.10$\,au expands
to reach a separation of about 0.126--0.145\,au for final masses for
the inner two stars of 1.4\,\MSun, which is consistent with the
observed twin BS 7782. In our simulations the eccentricity of the
inner binary grows to about $e \simeq 0.0028$.

With the accretion of mass, both stars in the inner binary also
accrete angular momentum.  By the end of the simulation the spins of
the two BSs are aligned along the orbital angular momentum axis with
an angle of $90.0^\circ$ for the primary star and $93.4^\circ$ for the
secondary star with respect to the argument of pericenter of the inner
orbit.  By the end of the simulations the spin of the primary is about
50.5 rotations per day, and 41.5 rotations per day for the secondary
star.
%% But by this time both stars have accreted
%% only a small portion of mass and we expect that the final spins of
%% both stars will be close to break-up (ignoring any spin-down from
%% magnetic braking effects).

@@

@@

As illustrated via SPH simulations, the mass transfer stream forms a
circumbinary disk, from which the inner binary stars accrete, driving
the inner binary toward a mass ratio close to unity.  Our simulations
indicate that the inner binary orbital separation can decrease or
expand depending on the details of the transfer of mass and angular
momentum.  More work is certainly needed in order to fully understand 
mass transfer in triples.
%Regretfully, we do not really understand the details of this
%process, but then this process is hardly understood in binary
%evolution also.
%NL: I edited the above to be a kinder to us.

@@

%NL2: Huh.  This is interesting.  Is the physical reason for it that
%apocenter for the secondary tends to be closer to the circumbinary
%disk?  I don't get why retrograde wins though, since naively I'd have
%thought prograde should lead to lower relative velocities between the
%secondary and the gas flow, and hence to higher accretion rates.
%What am I missing?  Is there momentum cancellation via colliding gas,
%shocks, etc. I am missing somewhere?
%% SPZ: I think that we do not have to add these extra simulations. I
%% found the result sufficiently confusing to check and check
%% again. Maybe we write another paper on these results, or a
%% follow-up on accretion in triples. But maybe it is best to stick to
%% the observed facts, and not try to explore parameter space in
%% extensio. What do you think, Nathan?  At the moment, this last
%% paragraph is a bit of an after though, but maybe it is worth
%% mentioning that our results are consistent with those of others...

In this paper, we propose a formation scenario for twin equal-mass
blue stragglers in tight binaries, as observed for WOCS 7782 in the
old OC NGC 188.  The proposed scenario involves mass transfer from an
evolved outer tertiary companion, part of this mass is accreted by the
inner binary via a circumbinary disk the rest escapes through the
second and third Lagrangian points in the potential of the triple
system.  Our scenario makes several predictions for the observed
properties of a hypothetical outer triple companion, now a WD.  These
are:

\item For the predicted outer tertiary orbit, the initial orbital
  period should lie between 220 days $\lesssim$ P$_{\rm out}$
  $\lesssim$ 1100 days, assuming initial masses for the inner binary
  components of $m_{\rm 1} = 1.1$ M$_{\odot}$ and $m_{\rm 2} = 0.9$
  M$_{\odot}$ and an initial outer tertiary mass of $m_{\rm 3} = 1.4
  $M$_{\odot}$.

\item The BSs in the inner binary should have roughly equal masses,
  independent of their initial masses.  This is because it is the
  lowest mass object that typically accretes the fastest, since its
  orbital velocity and distance relative to the circumbinary disk is
  typically the lowest
%% \citep[e.g.][]{1995MNRAS.277.1491K,1997MNRAS.285...33B,2000MNRAS.314...33B}.
  \citep[e.g.][]{2000MNRAS.314...33B,2012ApJ...749..118S,2017MNRAS.466.1170M}.
  The mass ratio of the inner binary, therefore grows to unity.  As a
  consequence, the initially lower mass MS star should accrete the
  most, and therefore be more enriched by accreted material.  This
  could be observable in the surface layers of a radiative star.  If
  the donor is an RGB star, the accretors will be enriched in mostly
  carbon, oxygen and helium, but if the donor is an AGB star the
  enrichment will be mostly in s-process elements.

\item We expect twin BSs in compact binaries formed from the mechanism
  proposed here to be more frequent in younger clusters with ages
  $\lesssim$ 4-6 Gyr.  This is because \textbf{stars in clusters} with a MS turn-off
  mass $\lesssim$ 1.2 M$_{\odot}$ have convective envelopes
  \citep[e.g.][]{1991ApJS...76...55I,2009pfer.book.....M}, and a
  radiative envelope for the donor in a mass transferring binary
  ensures stable accretion on to the accretor.  Note that part of the
  mass liberated from the triple system through the second and third
  Lagrangian points may eventually be accreted back onto the system.
  This could have interesting consequences for the enrichment of the
  low-mass white dwarf.

  %\simon{Can we quantify this? - NL: The stability of the accretion...?}

We emphasize, in closing, that the choice for the initial mass of the
outer tertiary may be rather critical.  Mass transfer in our proposed
scenario proceeds from the most massive tertiary to a binary of lower
total mass. This may result in an unstable phase of mass transfer, in
particular if the donor has a convective envelope
\citep[e.g.][]{2009pfer.book.....M}. A radiative envelope of the donor
ensures that the mass transfer will be maximally conservative, such
that the accretion stream will be maximally stable, accreting at a
stable and roughly constant rate \citep[e.g.][]{1991ApJS...76...55I}.
This stability regime may also be of interest for explaining very
massive twins, of $\apgt 20$\,\MSun\, which could be promising sources
for gravitational wave detectors once both twins evolve to a binary
black hole \citep{2016MNRAS.460.3545D}.

>>>>

We propose a formation mechanism for twin blue stragglers (BSs) in
compact binaries that involves mass transfer from an evolved outer
tertiary companion on to the inner binary via a circumbinary disk.  We
apply this scenario to the observed double BS binary WOCS 7782 in
the old open cluster NGC 188, and show that its observed properties
\textbf{can be} reproduced within the context of the proposed model.
Based on this model, we predict the following properties for twin BSs:
(1) For the outer tertiary orbit, the initial orbital period should
lie between 220 days $\lesssim$ P$_{\rm out}$ $\lesssim$ 1100 days,
assuming initial masses for the inner binary components of $m_{\rm 1}
= 1.1$ M$_{\odot}$ and $m_{\rm 2} =$ 0.9 M$_{\odot}$ and an outer
tertiary mass of $m_{\rm 3} = 1.4$ M$_{\odot}$.  \textbf{(2)} After Roche-lobe
overflow, the outer star turns into a white dwarf (WD) of mass 0.43 to
0.54\,\MSun. There is a correlation between the mass of this WD and
the outer orbital period: more massive WDs will be on wider orbits.
(3) The rotational axes of both BSs will be aligned with each other
and the orbital plane of the outer tertiary WD. (4) The BSs will have
roughly equal masses, independent of their initial masses (since the
lower mass star accretes the most).  The dominant accretor should,
therefore, be enriched more effectively by the accreted material.  As
a result, one of the BSs will appear to be more enriched by either He,
C and O or by s-process elements, depending on if the donor started to
overflow its Roche lobe on, respectively, the red giant or asymptotic
giant branch.  (5) Relative to old dense clusters with high-velocity
dispersions, twin BSs in close binaries formed from the proposed
mechanism should be more frequent in the Galactic field and younger
open clusters with ages $\lesssim$ 4-6 Gyr, since then the donor will
have a radiative envelope. (6) the orbit of the binary BS will have a
small semi-major axis (typically $\aplt 0.3$\,au) and be close to
circular ($e \aplt 0.2$).

@@

Even more curious, there exists in the old OC NGC 188 a
double BS binary, called WOCS 7782 \citep{geller08}.  The BS
population in NGC 188 has a bi-modal period-eccentricity distribution.
As discussed in \citet{2011MNRAS.410.2370L}, this could be hinting at
a triple origin for at least some subset of the total BS population.
As for WOCS 7782, \citet{2009Natur.462.1032M} observed a compact and
mildly eccentric (i.e., $e \sim 0.1$) binary star system with an
orbital period of $\sim$ 10 days hosting two roughly equal-mass BSs.  During a given binary-binary interaction, the probability
that not one but two direct (MS-MS) collisions will occur is less than
$10^{-2}$
\citep{1989AJ.....98..217L,2011MNRAS.410.2370L,2012MNRAS.425.2369L}.
Plus, binaries with collision products typically have relatively long
orbital periods \cite{2011Sci...334.1380F}. Dynamically, it is
difficult to form a short-period binary composed of two collision
products during a collisional interaction in a star cluster
\citep{2011MNRAS.410.2370L,2011Sci...334.1380F}, \textit{and the timescale for exchanging another BS into a pre-existing BS-MS or BS-WD binary is much longer than the expected BS lifetime (see \citet{2011MNRAS.410.2370L} and the end of Section~\ref{sect:dyn} below)}.  So, how did WOCS 7782 form?

@@

In our scenario, we start with a binary star with component masses
$m_{\rm 1}$ and $m_{\rm 2}$ that is orbited by a tertiary of mass
$m_3$. The inner and outer binary orbital semi-major axes are denoted
a$_{\rm in}$ and a$_{\rm out}$, respectively.  \textbf{For specificty,} we 
assume both orbits, the inner as well as the outer, to have negligible
eccentricity and low inclination, \textbf{which minimizes chaotic effects during the mass transfer process, facilitates more stable mass transfer and ultimately allows us to simulate our target system for longer while also maximizing the amount of mass transferred}.  These assumptions are also
supported by the population of observed low-mass triples
\citep{2010yCat..73890925T,2018ApJ...854...44M}.  This initial configuration for our
assumed formation scenario for WOCS 7782, described below, is
depicted in figure~\ref{fig:fig1}.

@@

Figure~\ref{fig:fig2} shows the parameter space in the P$_{\rm
  out}$-P$_{\rm in}$-plane for WOCS 7782.  We assume initial
component masses of $m_{\rm 1} = 1.1$ M$_{\rm \odot}$ and $m_{\rm 2} =
0.9$ M$_{\rm \odot}$ for the inner binary components, and $m_{\rm 3} =
1.4 $M$_{\rm \odot}$ for the outer tertiary.  We compare the
circularization radius to the semi-major axis of the inner binary, for
which we require $a_{\rm c} > a_{\rm in}$, after folding in all
constraints from the requirements for dynamical stability (listed in
the caption of figure~\ref{fig:fig2}), and the assumption of an outer tertiary that is Roche
lobe-filling \textbf{(see \citet{2014MNRAS.438.1909D} for more details)}.  
%\simon{Why do you consider it important to mention that the orbit is hard, and for what environment did you consider this? - NL: Deleted.}  
Note that the range of
plotted orbital periods P$_{\rm in}$ corresponding to a contact state
for the inner binary lies outside the range of plotted values for
$P_{\rm in}$ (for components with radii of 1 R$_{\odot}$),
since it does not contribute to constraining the outer orbital
properties.  The thick horizontal solid red line shows the allowed
range of outer semi-major axes, after folding in all of 
the aforementioned criteria.  These constraints result in a rather narrow
range of initial conditions for the outer orbit, namely 2.2 $\times$
10$^{2}$ days $\le$ P$_{\rm out}$ $\le$ 1.1 $\times$ 10$^3$ days, \textit{which also directly translates into constraints on the final outer tertiary orbit.}
%, for
%which the tertiary star overfills its Roche lobe, the triple is (at
%least initially) dynamically stable and the accretion stream forms a
%gaseous ring around the inner binary.

\textbf{Finally, we compute the timescales for our hypothesized triple to undergo a direct interaction with another single or binary star.  Using the same assumptions for the host cluster properties of NGC 188 outlined in Section 3.2 of \citet{2011MNRAS.410.2370L} (right-hand column), we find upon setting the single-triple (1+3) and binary-triple (2+3) timescales equal to the expected duration of the mass transfer phase (i.e., $\sim$ 1 Myr) critical outer orbital periods for triples of, respectively, 8.1 $\times$ 10$^8$ and 2.4 $\times$ 10$^9$ days.  These critical outer tertiary orbital periods, which correspond to the times for a \textit{specific} triple to undergo an interaction, correspond to 1+3 and 2+3 interaction times that are much longer than the maximum predicted outer orbital period of the hypothesized white dwarf tertiary in our scenario.  Hence, we do not expect the mass transfer process to be interrupted by either 1+3 and 2+3 interactions.}

@@

With the adopted parameters, we can also estimate the final orbital
period of the left-over core from the tertiary star after 
mass transfer.  The change in orbital separation due to
non-conservative mass transfer can be expressed in terms of the mass
of the outer star before and after mass transfer, i.e. $m_3$ and $m'_3$
respectively, the total mass in the inner binary before ($m_{\rm in}$)
and after accretion ($m'_{\rm in}$) and the amount of angular momentum
lost per unit mass $\eta \simeq 3$. Adopting the relation between the
orbital separation before mass transfer ($a$) and after mass transfer
($a'$) from \cite{1995A&A...296..691P}
\begin{equation}
  {a' \over a} = \left( {m_3 m_{\rm in} \over m'_3 m'_{\rm in}} \right)^{-2}
  \left( {m_3 + m_{\rm in} \over m'_3 + m'_{\rm in}} \right)^{2\eta + 1},
\end{equation}
we arrive at the top brown curve in
figure\,\ref{fig:tertiarymass_vs_size}. This curve provides a prediction for 
the current orbital separation of the WD around the twin
BS 7782. Tidal effects during mass \textbf{transfer}
have probably circularized the orbit, although some slight eccentricity
due to turbulent motion in the outer layers of the donor star may have
induced a small $e \ll 0.1$ eccentricity.

@@

We adopt initial masses of $m_{\rm 1} = 1.1$ M$_{\rm \odot}$ and
$m_{\rm 2} = 0.9$ M$_{\rm \odot}$ for the inner binary components, and
between $m_{\rm 3} = 1.2$ and $m_{\rm 3} = 1.4$ M$_{\rm \odot}$ for
the tertiary star.  We evolve the tertiary star using the MESA
stellar-evolution code \cite{2011ApJS..192....3P} to a radius of about
100\,\RSun\, and 150\,\RSun, at which point we assume it to overfill
it's Roche lobe (see red \textbf{squares} in
figure~\ref{fig:tertiarymass_vs_size}).  We perform calculations for an
inner orbital separation of $a_{\rm in} = 0.10$\,au, $a_{\rm in} =
0.15$\,au and $a_{\rm in} = 0.20$\,au.  In total we performed 12
calculations at a resolution of 40k SPH particles and 12 at 80k.

@@

With these parameters the orbital separation of the outer binary
becomes $\sim 250$ R$_{\rm \odot}$ for the 100 R$_{\rm \odot}$ donor star and about
430\,\RSun\, for the more evolved donor star.  We adopt the outer
orbit to be circular and in the plane of the inner binary.
%% We
%% performed several additional calculations for inclined orbits (choosing
%% a relative initial inclination $i = 9^\circ$, $i = 45^\circ$, $i =
%% 90^\circ$ and $i = 180^\circ$, but we wi
%% The low inclination is motivated by observed stellar triples composed
%% of low-mass stars to favour co-planar configurations
%% \citep{2018ApJ...854...44M}.
%% To test the effect of highly inclined
%% and retrogade orbits we include the $i = 90^\circ$ and $i = 180^\circ$
%% cases.
%%SPZ: I removed the notion of low-inclination, because we already said as such in the intro.
In figure~\ref{fig:topview_at_t0} we present a top view of the initial
conditions for one of these calculations.

@@

To test the \textbf{hypothesis that the} secondary in the inner binary
accretes more effectively than the primary star and to measure the
change to the inner orbit due to the Roche-lobe overflow of the outer
star, we perform a series of calculations in which we take the self
gravity and the hydrodynamical effects of the triple into account.
The results of these simulations are presented in
figure~\ref{fig:topview_at_t1000day} and
figure~\ref{fig:mass_vs_semimajor_axis}.  The first figure
(figure~\ref{fig:topview_at_t1000day}) shows the top view of the same
initial realization for which we presented the initial conditions in
figure~\ref{fig:topview_at_t0} but now at an age of 1091 days after
the onset of mass transfer. We add, to the left panel,
the equipotential surfaces in the orbital plane.

It is apparent that the mass transfer in the adopted triples leads to
a rather untidy evolution, since much of the donor mass is lost
through the second Lagrangian point to the right side of the donor
star in figure~\ref{fig:topview_at_t1000day}. A considerable amount of
mass is also lost through the third Lagrangian point (to the left of
the inner binary), although it is hard to actually \textbf{quantify the amount
of material lost}, becuase an appreciable fraction is expected to rain
back onto the triple system.  One remaining question is how much mass
is eventually ejected altogether from the triple system and is
therefore not accreted to any of the two inner stars. This value is
hard to estimate from the simulations, but an accretion efficiency of
$\apgt 0.6$ is necessary to make the scenario feasible. Over the time
scale for which we performed the calculations, this efficiency is
reached, but it is not clear how the system responds at later stages.

@@

In the right-hand panel in figure~\ref{fig:mass_vs_semimajor_axis} we
present the evolution of the orbit for the 1.4\,\MSun\, donor for
several initial orbits of the inner binary. A more massive donor
appears to be more effective in producing a twin BS with parameters
consistent with the observed system 7782. There is more mass available
in the envelope of the donor star, and the orbital evolution of the
inner binary matches better with the anticipated evolution \textbf{needed to reproduce the observed parameters of WOCS 7782.}  A more
massive donor may therefore have a lower accretion efficiency while
still accomodating the observed constraints.  The longer thermal time
scale of the stellar envelope of the higher-mass donor at the same
stellar radius eventually leads to a higher mass-transfer rate, and
therefore to a lower accretion efficiency. However, the larger mass
budget in the envelope appears to compensate.

The orbit of the inner binary expands in these cases as a result of
accretion onto the inner two stars. \textbf{In all} three cases for the
1.4\,\MSun\, donor presented in
figure~\ref{fig:mass_vs_semimajor_axis} the inner orbit expands at
about the same rate. Consequently, the inner binaries that start with
$a = 0.15$\,au and $a=0.20$\,au eventually become dynamically
unstable.  The binary with an initial separation of $0.10$\,au expands
to reach a separation of about 0.126--0.145\,au for final masses for
the inner two stars of 1.4\,\MSun, which is consistent with the
observed twin BS 7782. In our simulations the eccentricity of the
inner binary grows to about $e \simeq 0.0028$.

With the accretion of mass, both stars in the inner binary also
accrete angular momentum.  By the end of the simulation the spins of
the two BSs are aligned along the orbital angular momentum axis with
an angle of $90.0^\circ$ for the primary star and $93.4^\circ$ for the
secondary star with respect to the argument of pericenter of the inner
orbit.  \textbf{This supports our naive prediction that the spin angular momenta of the two stars in the inner binary should be more or less aligned post-mass transfer, due to the non-negligible amount of mass accreted.} By the end of the simulations the spin of the primary is about
50.5 rotations per day, and 41.5 rotations per day for the secondary
star.  \textbf{Such high spin rates immediately post-mass transfer are supported by other work (see, for example, \citet{2014MNRAS.438.1909D}), which could then subsequently spin down due to, for example, magnetic braking, bringing them closer to the actually observed spin rates.}
%% But by this time both stars have accreted
%% only a small portion of mass and we expect that the final spins of
%% both stars will be close to break-up (ignoring any spin-down from
%% magnetic braking effects).

@@

In this paper, we propose a formation scenario for twin equal-mass
blue stragglers in tight binaries, as observed for WOCS 7782 in the
old OC NGC 188.  The proposed scenario involves mass transfer from an
evolved outer tertiary companion\textbf{. Part} of this mass is accreted by the
inner binary via a circumbinary disk\textbf{, while} the rest escapes through the
second and third Lagrangian points in the potential of the triple
system.  Our scenario makes several predictions for the observed
properties of a hypothetical outer triple companion, now a WD.  These
are:

\item For the predicted outer tertiary orbit, the initial orbital
  period should lie between 220 days $\lesssim$ P$_{\rm out}$
  $\lesssim$ 1100 days, assuming initial masses for the inner binary
  components of $m_{\rm 1} = 1.1$ M$_{\odot}$ and $m_{\rm 2} = 0.9$
  M$_{\odot}$ and an initial outer tertiary mass of $m_{\rm 3} = 1.4          
  $M$_{\odot}$.  \textbf{These initial orbitl periods should match closely to the final periods we expected for the outer tertiary, since remaining within this range is needed to ensure Roche lobe overflow and we do not expect much subsequent orbital evolution post-mass transfer due to mass loss from stellar evolution.}

\item The BSs in the inner binary should have roughly equal masses,
  independent of their initial masses.  This is because it is the
  lowest mass object that typically accretes the fastest, since its
  orbital velocity and distance relative to the circumbinary disk is
  typically the lowest
%% \citep[e.g.][]{1995MNRAS.277.1491K,1997MNRAS.285...33B,2000MNRAS.314...33B\
%%}.                                                                            
  \citep[e.g.][]{2000MNRAS.314...33B,2012ApJ...749..118S,2017MNRAS.466.1170M}.
  The mass ratio of the inner binary, therefore grows to unity.

\item As a consequence of the above, the initially lower mass MS star should accrete the
  most, and therefore be more enriched by accreted material.  This
  could be observable in the surface layers of a radiative star.  If
  the donor is an RGB star, the accretors will be enriched in mostly
  carbon, oxygen and helium, but if the donor is an AGB star the
  enrichment will be mostly in s-process elements.  \textbf{Note that part of the
  mass liberated from the triple system through the second and third
  Lagrangian points may eventually be accreted back onto the system.
  This could also have interesting consequences for the enrichment of the
  low-mass white dwarf.}

\item We expect twin BSs in compact binaries formed from the mechanism
  proposed here to be more frequent in younger clusters with ages
  $\lesssim$ 4-6 Gyr.  This is because \textbf{stars in clusters} with a MS turn-off
  mass $\lesssim$ 1.2 M$_{\odot}$ have convective envelopes
  \citep[e.g.][]{1991ApJS...76...55I,2009pfer.book.....M}, and a
  radiative envelope for the donor in a mass transferring binary
  ensures stable accretion on to the accretor.

Finally, we emphasize that the choice for the initial mass of the
outer tertiary may be rather critical.  Mass transfer in our proposed
scenario proceeds from the most massive tertiary to a binary of lower
total mass. This may result in an unstable phase of mass transfer, in
particular if the donor has a convective envelope
\citep[e.g.][]{2009pfer.book.....M}. A radiative envelope of the donor
ensures that the mass transfer will be maximally conservative, such
that the accretion stream will be maximally stable, accreting at a
stable and roughly constant rate \citep[e.g.][]{1991ApJS...76...55I}.
This stability regime may also be of interest for explaining very
massive twins, of $\apgt 20$\,\MSun\, which could be promising sources
for gravitational wave detectors once both twins evolve to a binary
black hole \citep{2016MNRAS.460.3545D}.

@@

As illustrated via SPH simulations, the mass transfer stream forms a
circumbinary disk, from which the inner binary stars accrete, driving
the inner binary toward a mass ratio close to unity.  Our simulations
indicate that the inner binary orbital separation can decrease or
expand depending on the details of the transfer of mass and angular
momentum.  More work is certainly needed in order to fully understand
mass transfer in triples.

@@
